Data Cleaning logbook

### 2018-06-30 docs/paper/Villasenor-Derbez.tex
Added comments by coauthors

\def\keyFont{\fontsize{8}{11}\helveticabold }
\def\firstAuthorLast{Villaseñor-Derbez {et~al.}}
\def\Authors{Juan Carlos Villaseñor-Derbez\(^{1,*}\), Eréndira
Aceves-Bueno\(^{1,*}\), Álvin Suarez-Castillo\(^{2}\), Stuart
Fulton\(^{2}\), Jorge Torre\(^{2}\)}
% Affiliations should be keyed to the author's name with superscript numbers and be listed as follows: Laboratory, Institute, Department, Organization, City, State abbreviation (USA, Canada, Australia), and Country (without detailed address information such as city zip codes or street names).
% If one of the authors has a change of address, list the new address below the correspondence details using a superscript symbol and use the same symbol to indicate the author in the author list.
\def\Address{\(^{1}\)Bren School of Environmental Science and Management, University
of California, Santa Barbara, Santa Barbara, CA,
USA\newline \(^{2}\)Comunidad y Biodiversidad A.C., Guaymas, Mexico}
% The Corresponding Author should be marked with an asterisk
% Provide the exact contact address (this time including street name and city zip code) and email of the corresponding author
\def\corrAuthor{Juan Carlos Villaseñor-Derbez, Bren Hall, University of California,
Santa Barbara, Santa Barbara, CA, 93106}

@@

\textbf{Titulos alternativos}

``Management and effectiveness of community-based marine reserves in
small-scale fisheries''

``Community-based marine reserves in small-scale fisheries''

``Effectiveness of community-based marine reserves: lessons for their
management and implementation'' (no me encanta)

Algun otro?

\textbf{Last update}: 2018-06-11

@@

Marine ecosystems around the world sustain significant impacts due to
overfishing and unsustainable fishing practices
\citep{halpern_2008-dK,worm_2006-IB,pauly_2005-qV}. A common approach to
manage the spatial distribution of fishing effort to recover stocks and
preserve biodiversity is through the implementation of marine reserves.
These areas allow bounded populations to recover by limiting all
extractive activities \citep{halpern_2002}.

The science of marine reserves has largely focused on understanding the
ecological effects of these areas, which include increased biomass,
richness, and densities of organisms within the protected regions,
climate change mitigation, and protection from environmental variability
\citep{lester_2009-Ks,giakoumi_2017-V2,sala_2017-69,roberts_2017-J9,micheli_2012-EU}.
Modelling studies show that fishery benefits of marine reserves depend
on initial stock status and the management under which the fishery
operates, as well as reserve size and the amount of larvae exported from
these \citep{hilborn_2006,krueck_2017-J1}. Other research has focused on
the relationship between socioeconomic and governance structures and
their relationship to reserve effectiveness
\citep{halpern_2013,lpezangarita_2014,mascia_2017-m_}. However, few
studies simultaneously evaluate reserves from all these perspectives
(\emph{e.g.} \citet{lpezangarita_2014}).

Implementation of marine reserves can be done by following one of three
main approaches in terms of implementation. We describe these in the
context of Mexican MRs, but argue that at least the first two apply
elsewhere. Marine reserves are usually implemented and managed by a
government agency, in this case the National Commission of Natural
Protected Areas (\emph{Comisión Nacional de Áreas Naturales Protegidas},
CONANP). While CONANP has made efforts to incorporate stakeholders in
the design and implementation phases, these are still characterized by a
heavy top-down process. A second approach is the implementation of
community-based marine reserves, usually placed within areas of
exclusive access (\emph{i.e.} TURFs). Community-based spatial closures
occur in other places, like the \emph{kapu} or \emph{ra'ui} areas in the
Pacific Islands \citep{bohnsack_2004,johannes_2002}. This bottom-up
approach increases compliance and self-enforcement
\citep{gelcich_2015-Gw,espinosaromero_2014-PY,beger_2004-Y8}. However,
without legal recognition these are difficult to enforce and fishers
rely on the exclusive access granted by the TURF. In an effort to bridge
this normative gap, Civil Society Organizations (CSOs) served as the
link between fishers and government, and set out to create a legal
framework that solve this governance issue. In 2014, a new norm was
created, allowing fishers to request the legal recognition of a
community-based reserve under the name of ``Fishing Refugia''
\citep{nom}. These can be implemented as temporal or partial reserves,
which can protect one, some, or all resources within them. Since then,
\textbf{41} of community-based marine reserves along the Pacific, Gulf
of California, and Mexican Caribbean coastlines have gained legal
recognition, but no formal evaluation of their effectiveness has taken
place.

This work combines causal inference techniques and the social-ecological
systems framework to provide a holistic evaluation of community-based
marine reserves in three coastal communities in Mexico. The objective of
this work is twofold. First, provide a triple bottom line evaluation of
the effectiveness of community-based marine reserves that can inform
similar processes in other countries. And second, perform the first
formal evaluation of Fishing Refugia in Mexico and identify areas where
improvement or adjustment might result in increased effectiveness. On
both cases, we draw from the lessons learned and provide management
recommendations to maximize the effectiveness of community-based marine
reserves in small-scale fisheries.

@@

We evaluate community-based marine reserves from three coastal
communities located in the Pacific coast of Baja California and the
Mexican Caribbean (Fig \ref{fig:map}). All communities are organized as
fishing cooperatives that hold Territorial Use Rights for Fisheries
(TURFs). Isla Natividad lies west of the Baja California Peninsula (Fig
\ref{fig:map}B), where kelp forests and rocky reefs are the predominant
habitats. The island is home to the \emph{Buzos y Pescadores de la Baja
California} fishing cooperative, whose main resource by value is the
spiny lobster (\emph{Panulirus interruptus}). Other resources like
finfish, sea cucumber, red sea urchin, snail, and abalone are also
important sources of income. In 2006, the community decided to implement
two community-based marine reserves within their fishing grounds to
protect commercially important invertebrate species (mainly lobster and
abalone). These reserves obtained legal recognition in 2018, but were
well enforced since their implementation in 2006.

The other two communities are Maria Elena and Punta Herrero, which are
located in the Yucatan Peninsula, where coral reefs and mangroves are
the representative coastal ecosystems( Fig \ref{fig:map}C). Maria Elena
is a fishing camp --visited intermittently during the fishing season--
belonging to the Cozumel fishing cooperative (\emph{SCPP Cozumel});
Punta Herrero is home to the \emph{SCPP José María Azcorra} cooperative.
Their main fishery is the Caribbean spiny lobster (\emph{Panulirus
argus}), but they also target finfish in the off-season. Maria Elena and
Punta Herrero established eight marine reserves in 2012, and four marine
reserves in 2013, respectively. All these reserves are legally
recognized as Fishing Refugia since their creation.

@@

We use three main sources of information to evaluate these reserves
across the ecological, socioeconomic, and governance dimensions.
Ecological data come from the annual ecological monitoring of reserve
and control areas, carried out by members from each community and
personnel from the Mexican CSO \emph{Comunidad y Biodiversidad}
(\href{www.cobi.org.mx}{COBI}). Trained divers record richness and
abundances of fish and invertebrate species in the reserves and control
sites. Size structures are also collected during fish surveys. We define
control sites as regions with habitat characteristics similar to the
corresponding reserves, and that presumably had a similar probability of
being selected as reserves during the design phase. We focus our
evaluation on sites where data are available for reserve and control
sites, before and after the implementation of the reserve. This provides
us with a Before-After-Control-Impact (\emph{i.e.} BACI) sampling design
that allows us to capture and control for temporal and spatial dynamics
\citep{depalma_2018,ferraro_2006-oW}. BACI designs and causal inference
techniques have proven effective to evaluate marine reserves, as they
allow us to causally attribute observed changes to the intervention
\citep{moland_2013-VP,Villasenor-Derbez_2018}. All sites were surveyed
annually, and at least once before implementation of the reserves. Table
\ref{table:com_sum} shows a summary of the reserves included in this
study.

\begin{figure}
\centering
\includegraphics{Villasenor-Derbez_files/figure-latex/unnamed-chunk-1-1.pdf}
\caption{\label{fig:unnamed-chunk-1}\label{fig:map}Location of the three
coastal communities studied (A). Isla Natividad (B) is located off the
Baja California Peninsula, Maria Elena (C) and Punta Herrero (D) are
located in the Yucatan Peninsula. Blue polygons represent the TURFs, and
red polygons the marine reserves.}
\end{figure}

@@

Indicators showed ambiguous responses through time for each reserve.
Figure \ref{fig:indicators}A shows positive effect sizes for lobster
densities in Isla Natividad and Punta Herrero during the first years,
but the effect is eroded through time. These effects are in the order of
0.2 extra organisms \(\mathrm{m}^{-2}\) but are not significantly
different from zero (\(p > 0.05\)). Lobster densities were only
significantly positive for Isla Natividad on the sixth year (\emph{i.e.}
2012; \(p < 0.05\)), a year after the hypoxia events described by
\citet{micheli_2012-EU} caused mass mortality of organisms. Likewise, no
changes were detected in fish biomass or invertebrate and fish densities
(\ref{fig:indicators}B-D), where effect sizes oscillated around zero
without clear trends. Full tables with model coefficients are presented
in the supplementary materials (\textbf{S2 Table}, \textbf{S3 Table},
\textbf{S4 Table}).

@@

Literature shows that age, isolation, and enforcement are important
factors that influence reserve effectiveness \citep{edgar_2014-UO}. Isla
Natividad has the oldest reserve, and all communities are fairly
isolated, and have a well-established community-based enforcement
system. With these characteristics, one would expect the reserves to be
effective. However, another key condition for effectiveness is reserve
is size \citep{edgar_2014-UO}, and the lack of effectiveness can perhaps
be attributed to reserves being too small. Previous research has shown
that reserves in Isla Natividad yield fishery benefits for the abalone
fishery \citep{rossetto_2015-V0}. Abalone are less mobile than lobsters,
and perhaps the reserves provide enough protection to these sesile
invertebrates, but not lobsters. Maria Elena and Punta Herrero are
relatively young reserves, and it is known that community-based marine
reserves in tropical ecosystems may take up to six years to show a
spillover effect \citep{dasilva_2015-zX}. Work by \citet{ayer_2018}
shows these communities support the implementation of marine reserves.
Nevertheless, fishers may favor implementation of reserves that pose low
fishing costs either because of their location or size. The economic
data support our claim that reserves are small, as neither landings nor
revenues showed the expected short-term costs associated to the first
years of reserve implementation \citep{ovando_2016-Wg}.

Even if reserves had appropriate sizes, there are other plausible
explanations for the observed lack of effectiveness. Marine reserves are
only likely to provide fisheries benefits if initial population sizes
are low and the fishery is poorly managed \citep{hilborn_2006}. However,
both lobster fisheries were, at some point, certified by the Marine
Stewardship Council \citep{prezramrez_2016-J1}. Additionally, lobster
fisheries are managed via species-specific minimum catch sizes, seasonal
closures, protection of ``berried'' females, and escapement windows
where traps are allowed \cite{dof_website_1993}. It is uncertain whether
that such a well-managed fishery will experience additional benefits
from marine reserves.

While reserves fail to provide fishery benefits, there are a number of
additional ecological, fisheries, and social benefits. Marine reserves
provide protection to a wider range of species and vulnerable habitat,
like coral reefs. These sites can serve as an insurance against
environmental shocks or mistakes in fisheries management
\citep{hilborn_2004,hilborn_2006,micheli_2012-EU}. Embarking in a marine
conservation project can bring the community together, which promotes
social cohesion and builds social capital. Furthermore, showing
commitment to marine conservation allows fishers to have greater
bargaining power and leverage over fisheries management.

@@

JCVD CONACyT (Beca de Posgrados en el extranjero, CVU 669403) + LAFF +
El proyecto de COBI que haya pagado esto

\textbf{ASC}

\textbf{SF}

\textbf{JT}

\textbf{Que fuentes de financiamiento incluyo par ustedes?}

>>>>

\def\keyFont{\fontsize{8}{11}\helveticabold }
\def\firstAuthorLast{Villaseñor-Derbez {et~al.}}
\def\Authors{Juan Carlos Villaseñor-Derbez\(^{1,*}\), Eréndira
Aceves-Bueno\(^{1,*}\), Álvin Suarez\(^{2}\), Stuart Fulton\(^{2}\),
Jorge Torre\(^{2}\)}
% Affiliations should be keyed to the author's name with superscript numbers and be listed as follows: Laboratory, Institute, Department, Organization, City, State abbreviation (USA, Canada, Australia), and Country (without detailed address information such as city zip codes or street names).
% If one of the authors has a change of address, list the new address below the correspondence details using a superscript symbol and use the same symbol to indicate the author in the author list.
\def\Address{\(^{1}\)Bren School of Environmental Science and Management, University
of California, Santa Barbara, Santa Barbara, CA,
USA\newline \(^{2}\)Comunidad y Biodiversidad A.C., Guaymas, Mexico}
% The Corresponding Author should be marked with an asterisk
% Provide the exact contact address (this time including street name and city zip code) and email of the corresponding author
\def\corrAuthor{Juan Carlos Villaseñor-Derbez, Bren Hall, University of California,
Santa Barbara, Santa Barbara, CA, 93106}

@@

\textbf{Last update}: 2018-06-29

@@

Marine ecosystems around the world sustain significant impacts due to
overfishing and unsustainable fishing practices
\citep{halpern_2008-dK,worm_2006-IB,pauly_2005-qV}. In particular,
artisanal fisheries face great challenges since they tend to be hard to
monitor and enforce \citep{costello_2012}. Recent research shows that
combining Territorial Use Rights for Fisheries (TURFs) with no-take
marine reserves (MR) can greatly improve the performance of coastal
fisheries and the health of the local resouces
\citep{costello_2010-Ix,lester_2017}. Commonly known as TURF-Reserves,
these systems increase the benefits of spatial access rights allowing
the maintainance of healthy resources
\citep{afflerbach_2014-HP,lester_2017}.

Marine reserves allow bounded populations to recover by limiting all
extractive activities \citep{halpern_2002}, while the TURF controls the
harvest levels outside the MPA. Although in theory these systems are
successful \citep{costello_2010-Ix}, little empirical evidence evidence
exists of their effectiveness and the drivers of their success
\citep{afflerbach_2014-HP,lester_2017}. The performance of these systems
depends on how environmental and social factors work combined. The
science of marine reserves has largely focused on understanding the
ecological effects of these areas, which include increased biomass,
richness, and densities of organisms within the protected regions,
climate change mitigation, and protection from environmental variability
\citep{lester_2009-Ks,giakoumi_2017-V2,sala_2017-69,roberts_2017-J9,micheli_2012-EU}.
Modelling studies show that fishery benefits of marine reserves depend
on initial stock status and the management under which the fishery
operates, as well as reserve size and the amount of larvae exported from
these \citep{hilborn_2006,krueck_2017-J1}. Other research has focused on
the relationship between socioeconomic and governance structures and
their relationship to reserve effectiveness
\citep{halpern_2013,lpezangarita_2014,mascia_2017-m_}. However, to our
knowledge, no studies exist that evaluate TURF-reserves from both a
social and ecological perspective.

In Mexico many of these marine reserves are created as community-based
marine reserves. Community-based spatial closures occur in other places,
like the \emph{kapu} or \emph{ra'ui} areas in the Pacific Islands
\citep{bohnsack_2004,johannes_2002}. This bottom-up approach increases
compliance and self-enforcement
\citep{gelcich_2015-Gw,espinosaromero_2014-PY,beger_2004-Y8}. However,
without legal recognition these are difficult to enforce and fishers
rely on the exclusive access granted by the TURF. In an effort to bridge
this normative gap, Civil Society Organizations (CSOs) served as the
link between fishers and government, and set out to create a legal
framework that solve this governance issue. In 2014, a new norm was
created, allowing fishers to request the legal recognition of a
community-based reserve under the name of ``Fish Refuge'' \citep{nom}.
These can be implemented as temporal or partial reserves, which can
protect one, some, or all resources within them. Since then, 45 of
community-based marine reserves along the Pacific, Gulf of California,
and Mexican Caribbean coastlines have gained legal recognition, but
their effectiveness has not been reported in the scientific literature.

This work combines causal inference techniques and the social-ecological
systems framework to provide a holistic evaluation of community-based
marine reserves in three coastal communities in Mexico. The objective of
this work is twofold. First, provide a triple bottom line evaluation of
the effectiveness of community-based marine reserves that can inform
similar processes in other countries. And second, evaluate the
effectiveness of TURF-reserves established as Fishing Refugia in Mexico
to identify areas where improvement or adjustment might result in
increased effectiveness. On both cases, we draw from the lessons learned
and provide management recommendations to maximize the effectiveness of
community-based marine reserves in small-scale fisheries.

@@

We evaluate three TURF-reserves in Mexico (Fig \ref{fig:map}A). The
first one was created by the \emph{Buzos y Pescadores de la Baja
California} fishing cooperative, located in Isla Natividad in the Baja
Peninsula (Fig \ref{fig:map}B). The main fishery in the island is the
spiny lobster (\emph{Panulirus interruptus}), but other resources like
finfish, sea cucumber, read sea urchin, snail, and abalone are also an
important source of income. In 2006, the community decided to implement
two marine reserves within their fishing grounds to protect commercially
important invertebrate species; mainly lobster and abalone. The reserves
obtained legal recognition in 2018, but have been well enforced since
their implementation.

The other two TURF-reserves are located in Maria Elena and Punta
Herrero, in the Yucatan Peninsula (Fig \ref{fig:map}C). Maria Elena is a
fishing camp --visited intermittently during the fishing season--
belonging to the Cozumel fishing cooperative (\emph{SCPP Cozumel});
Punta Herrero is home to the \emph{SCPP José María Azcorra} cooperative.
Their main fishery is the Caribbean spiny lobster (\emph{Panulirus
argus}), but they also target finfish in the off-season. Maria Elena and
Punta Herrero established eight marine reserves in 2012, and four marine
reserves in 2013, respectively. All these reserves are legally
recognized as Fishing Refugia since their creation.

@@

We use three main sources of information to evaluate these reserves
across the ecological, socioeconomic, and governance dimensions.
Ecological data come from the annual ecological monitoring of reserve
and control areas, carried out by members from each community and
personnel from the Mexican CSO \emph{Comunidad y Biodiversidad}
(\href{www.cobi.org.mx}{COBI}). Trained divers record richness and
abundances of fish and invertebrate species in the reserves and control
sites \citep{fulton_2018}. Size structures are also collected during
fish surveys. We define control sites as regions with habitat
characteristics similar to the corresponding reserves, and that
presumably had a similar probability of being selected as reserves
during the design phase. We focus our evaluation on sites where data are
available for reserve and control sites, before and after the
implementation of the reserve. This provides us with a
Before-After-Control-Impact (\emph{i.e.} BACI) sampling design that
allows us to capture and control for temporal and spatial dynamics
\citep{depalma_2018,ferraro_2006-oW}. BACI designs and causal inference
techniques have proven effective to evaluate marine reserves, as they
allow us to causally attribute observed changes to the intervention
\citep{moland_2013-VP,Villasenor-Derbez_2018}. All sites were surveyed
annually, and at least once before implementation of the reserves. Table
\ref{table:com_sum} shows a summary of the reserves included in this
study.

\begin{figure}
\centering
\includegraphics{Villasenor-Derbez_files/figure-latex/unnamed-chunk-1-1.pdf}
\caption{\label{fig:unnamed-chunk-1}\label{fig:map}Location of the three
coastal communities studied (A). Isla Natividad (B) is located off the
Baja California Peninsula, Maria Elena and Punta Herrero (C) are located
in the Yucatan Peninsula. Blue polygons represent the TURFs, and red
polygons the marine reserves.}
\end{figure}

@@

Indicators showed ambiguous responses through time for each reserve.
Figure \ref{fig:indicators}A shows positive effect sizes for lobster
densities in Isla Natividad and Punta Herrero during the first years,
but the effect is eroded through time. In the case of Maria Elena,
positive changes were observed in the third and forth year. These
effects are in the order of 0.2 extra organisms \(\mathrm{m}^{-2}\) for
Isla Natividad and Punta Herrero, and 0.01 organisms \(\mathrm{m}^{-2}\)
for Maria Elena, but are not significantly different from zero
(\(p > 0.05\)). The rapid increase observed for changes in lobster
densities for Isla Natividad on the sixth year (\emph{i.e.} 2012) occur
a year after the hypoxia events described by \citet{micheli_2012-EU}
caused mass mortality of organisms. Likewise, no changes were detected
in fish biomass or invertebrate and fish densities
(\ref{fig:indicators}B-D), where effect sizes oscillated around zero
without clear trends. Full tables with model coefficients are presented
in the supplementary materials (\textbf{S2 Table}, \textbf{S3 Table},
\textbf{S4 Table}).

@@

Literature shows that age and enforcement are important factors that
influence reserve effectiveness \citep{edgar_2014-UO}. Isla Natividad
has the oldest reserve, and our SES analysis suggests that all
communities have a well-established community-based enforcement system.
With these characteristics, one would expect the reserves to be
effective. Maria Elena and Punta Herrero are relatively young reserves
(\emph{i.e.} \textless{} 5 years old); other community-based marine
reserves in tropical ecosystems may take up to six years to show a
spillover effect \citep{dasilva_2015-zX}.

Another key condition for effectiveness is reserve size
\citep{edgar_2014-UO}, and the lack of effectiveness can perhaps be
attributed to reserves being too small. Previous research has shown that
reserves in Isla Natividad yield fishery benefits for the abalone
fishery \citep{rossetto_2015-V0}. Abalone are less mobile than lobsters,
and perhaps the reserves provide enough protection to these sesile
invertebrates, but not lobsters. Design principles developed by
\citet{green_2017} for marine reserves in the Caribbean state that
reserves ``should be more than twice the size of the home range of
adults and juveniles'', and suggest that reserves seeking to protect
spiny lobsters should have at least 14 km across. Furthermore, may favor
implementation of reserves that pose low fishing costs due to their
small size or location. Our analysis of economic data supports this, as
neither landings nor revenues showed the expected short-term costs
associated to the first years of reserve implementation
\citep{ovando_2016-Wg}. Small reserves can serve as a way to
self-regulate the spatial distribution of fishing effort

Even if reserves had appropriate sizes and were placed in optimal
locations, there are other plausible explanations for the observed
patterns. For instance, marine reserves are only likely to provide
fisheries benefits if initial population sizes are low and the fishery
is poorly managed \citep{hilborn_2006}. Both lobster fisheries were, at
some point, certified by the Marine Stewardship Council
\citep{prezramrez_2016-J1}. Additionally, lobster fisheries are managed
via species-specific minimum catch sizes, seasonal closures, protection
of ``berried'' females, and escapement windows where traps are allowed
\cite{dof_website_1993}. It is uncertain whether such a well-managed
fishery will experience additional benefits from marine reserves.

While reserves fail to provide fishery benefits, there are a number of
additional ecological, fisheries, and social benefits. Marine reserves
provide protection to a wider range of species and vulnerable habitat,
like coral reefs. These sites can serve as an insurance against
environmental shocks or mistakes in fisheries management
\citep{hilborn_2004,hilborn_2006,micheli_2012-EU}. Self-regulation of
fishing effort (\emph{i.e.} reduction in harvest) can serve as a way to
compensate for future declines associated to environmental variation
\citep{finkbeiner_2018}. Embarking in a marine conservation project can
bring the community together, which promotes social cohesion and builds
social capital. Furthermore, showing commitment to marine conservation
allows fishers to have greater bargaining power and leverage over
fisheries management.

@@

JCVD: CONACyT (Beca de Posgrados en el extranjero, CVU 669403) and the
Latin American Fisheries Fellowship.

AS, SF and JT Walton Family Foundation, Summit Foundation, and Oak
Foundation.

### 2018-06-30 scripts/data_cleaning_logbook.Rmd
Updated data cleaning
Added comments by coauthors

```{r}
invert <- read.csv(here::here("raw_data", "invert_data.csv"), strip.white = T, stringsAsFactors = F) %>% 
  janitor::clean_names() %>% 
  filter(comunidad %in% c("Isla Natividad", "Maria Elena", "Punta Herrero")) %>% 
  filter(!is.na(rc),
         rc %in% c("La Plana / Las Cuevas - La Dulce / Babencho",
                   "Cabezo - Cabezo (Control)",
                   "Manchon - Manchon (Control)",
                   "El Faro - El Faro (Control)")) %>% 
  filter(!(rc %in% c("El Faro - El Faro (Control)") & ano == 2012)) %>% 
  filter(between(abundancia, 0.1, 100)) %>% 
  group_by(comunidad, ano, rc, sitio, zona, transecto, generoespecie) %>% 
  summarize(abundancia = sum(abundancia)) %>% 
  ungroup() %>% 
  mutate(year = case_when(comunidad == "Isla Natividad" ~ ano - 2006,
                         comunidad == "Maria Elena" ~ ano - 2012,
                         TRUE ~ ano - 2013)) %>% 
  select(comunidad, ano, year, rc, sitio, zona, transecto, generoespecie, abundancia) %>% 
  arrange()
```

@@

```{r}
species_bio <- MPAtools::species_bio

@@

species_bio %<>%
  select(GeneroEspecie, a, b) %>% 
  rbind(FB_update) %>% 
  rbind(GenusLevelBio) %>% 
  janitor::clean_names()
```

@@

```{r}
fish <- read.csv(here::here("raw_data", "fish_data.csv"), strip.white = T, stringsAsFactors = F) %>% 
  janitor::clean_names() %>% 
  filter(comunidad %in% c("Isla Natividad", "Maria Elena", "Punta Herrero")) %>% 
  filter(!is.na(rc),
         rc %in% c("La Plana / Las Cuevas - La Dulce / Babencho",
                   "Cabezo - Cabezo (Control)",
                   "Manchon - Manchon (Control)",
                   "El Faro - El Faro (Control)")) %>% 
  filter(!(rc %in% c("El Faro - El Faro (Control)") & ano == 2012)) %>% 
  filter(abundancia > 0) %>% 
  mutate(generoespecie = case_when(generoespecie == "Cephalopholis cruentatus" ~ "Cephalopholis cruentata",
                                   generoespecie == "Haemulon macrostomun" ~ "Haemulon macrostomum",
                                   generoespecie == "Xcochin spp" ~ "Balistes capriscus",
                                   generoespecie == "Pterois volitans/miles" ~ "Pterois volitans",
                                   generoespecie == "Mycteroperca spp" ~ "Mycteroperca",
                                   generoespecie == "Chromis spp" ~ "Chromis",
                                   generoespecie == "Damisela spp" ~ "Stegastes",
                                   generoespecie == "Haemulon spp" ~ "Haemulon",
                                   generoespecie == "Lancero spp" ~ "Acanthurus",
                                   generoespecie == "Kyphosus sectatrix/incisor" ~ "Kyphosus",
                                   generoespecie == "Sebastes spp" ~ "Sebastes",
                                   generoespecie == "Loro spp" ~ "Scarus",
                                   generoespecie == "Lutjanus spp" ~ "Lutjanus",
                                   generoespecie == "Mariposa spp" ~ "Chaetodon",
                                   generoespecie == "Pluma spp" ~ "Calamus pennatula",
                                   generoespecie == "Sparisoma aurolineatum" ~ "Sparisoma aurofrenatum",
                                   generoespecie == "Vieja spp" ~ "Bodianus rufus",
                                   TRUE ~ generoespecie)) %>% 
  left_join(species_bio, by = c("generoespecie")) %>% 
  group_by(comunidad, ano, rc, sitio, zona, transecto, generoespecie, talla, a,b) %>% 
  summarize(abundancia = sum(abundancia)) %>% 
  ungroup() %>% 
  mutate(year = case_when(comunidad == "Isla Natividad" ~ ano - 2006,
                         comunidad == "Maria Elena" ~ ano - 2012,
                         TRUE ~ ano - 2013)) %>% 
  select(comunidad, ano, year, rc, sitio, zona, transecto, generoespecie, talla, abundancia, a, b) %>% 
  arrange()
```

>>>>

```{r}
invert <- read.csv(here::here("raw_data", "invert_data.csv"), strip.white = T, stringsAsFactors = F) %>% 
  janitor::clean_names() %>% 
  rename(generoespecie = genero_especie) %>% 
  filter(comunidad %in% c("Isla Natividad", "Maria Elena", "Punta Herrero")) %>% 
  filter(!is.na(rc),
         rc %in% c("La Plana / Las Cuevas - La Dulce / Babencho",
                   "Cabezo - Cabezo (Control)",
                   "Manchon - Manchon (Control)",
                   "El Faro - El Faro (Control)")) %>% 
  filter(!(rc %in% c("El Faro - El Faro (Control)") & ano == 2012)) %>% 
  filter(between(abundancia, 0.1, 100)) %>% 
  group_by(comunidad, ano, rc, sitio, zona, transecto, generoespecie) %>% 
  summarize(abundancia = sum(abundancia)) %>% 
  ungroup() %>% 
  mutate(year = case_when(comunidad == "Isla Natividad" ~ ano - 2006,
                         comunidad == "Maria Elena" ~ ano - 2012,
                         TRUE ~ ano - 2013)) %>% 
  select(comunidad, ano, year, rc, sitio, zona, transecto, generoespecie, abundancia) %>% 
  arrange()
```

@@

```{r}
species_bio <- MPAtools::species_bio %>% 
  mutate(Genus = as.character(str_split(GeneroEspecie, pattern = " ", simplify = T)[,1]),
         Genus = str_replace(Genus, " ", ""))

@@

species_bio %<>%
  select(GeneroEspecie, a, b) %>% 
  rbind(FB_update) %>% 
  rbind(GenusLevelBio) %>% 
  janitor::clean_names() %>% 
    rename(generoespecie = genero_especie)
```

@@

```{r}
fish <- read.csv(here::here("raw_data", "fish_data.csv"), strip.white = T, stringsAsFactors = F) %>% 
  janitor::clean_names() %>% 
  rename(generoespecie = genero_especie) %>% 
  filter(comunidad %in% c("Isla Natividad", "Maria Elena", "Punta Herrero")) %>% 
  filter(!is.na(rc),
         rc %in% c("La Plana / Las Cuevas - La Dulce / Babencho",
                   "Cabezo - Cabezo (Control)",
                   "Manchon - Manchon (Control)",
                   "El Faro - El Faro (Control)")) %>% 
  filter(!(rc %in% c("El Faro - El Faro (Control)") & ano == 2012)) %>% 
  filter(abundancia > 0) %>% 
  mutate(generoespecie = case_when(generoespecie == "Cephalopholis cruentatus" ~ "Cephalopholis cruentata",
                                   generoespecie == "Haemulon macrostomun" ~ "Haemulon macrostomum",
                                   generoespecie == "Xcochin spp" ~ "Balistes capriscus",
                                   generoespecie == "Pterois volitans/miles" ~ "Pterois volitans",
                                   generoespecie == "Mycteroperca spp" ~ "Mycteroperca",
                                   generoespecie == "Chromis spp" ~ "Chromis",
                                   generoespecie == "Damisela spp" ~ "Stegastes",
                                   generoespecie == "Haemulon spp" ~ "Haemulon",
                                   generoespecie == "Lancero spp" ~ "Acanthurus",
                                   generoespecie == "Kyphosus sectatrix/incisor" ~ "Kyphosus",
                                   generoespecie == "Sebastes spp" ~ "Sebastes",
                                   generoespecie == "Loro spp" ~ "Scarus",
                                   generoespecie == "Lutjanus spp" ~ "Lutjanus",
                                   generoespecie == "Mariposa spp" ~ "Chaetodon",
                                   generoespecie == "Pluma spp" ~ "Calamus pennatula",
                                   generoespecie == "Sparisoma aurolineatum" ~ "Sparisoma aurofrenatum",
                                   generoespecie == "Vieja spp" ~ "Bodianus rufus",
                                   TRUE ~ generoespecie)) %>% 
  left_join(species_bio, by = c("generoespecie")) %>% 
  group_by(comunidad, ano, rc, sitio, zona, transecto, generoespecie, talla, a,b) %>% 
  summarize(abundancia = sum(abundancia)) %>% 
  ungroup() %>% 
  mutate(year = case_when(comunidad == "Isla Natividad" ~ ano - 2006,
                         comunidad == "Maria Elena" ~ ano - 2012,
                         TRUE ~ ano - 2013)) %>% 
  select(comunidad, ano, year, rc, sitio, zona, transecto, generoespecie, talla, abundancia, a, b) %>% 
  arrange() %>% 
  drop_na(a, b)
```
